---
title: "Appendix 1: Toronto Data Models"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F, cache=T)
```

```{r message=F}
library(sf)
library(INLA)
library(tidyverse)
library(spdep)
```

```{r include=F}
data_csv = read_csv("data/data_clean.csv")
data = st_read("data/data.shp")
names(data) = c(names(data_csv), "geometry")
```

```{r}
# Filter out missing data
data = data %>% drop_na()
```

```{r}
# specify priors
prior = list(
  prec = list(
    prior = "pc.prec",
    param = c(0.5 / 0.31, 0.01)),
  phi = list(
    prior = "pc",
    param = c(0.5, 0.5))
  )
```

```{r message=F}
# Compute adjacency matrix and graph
nb = poly2nb(data)
nb2INLA("map.adj", nb)
g = inla.read.graph(filename = "map.adj")

# Neighbourhood index 
spatial = 1:nrow(data)
```

## Violent crime models

The following models have violent crime as the response, and have the same socio-economic covariates, but with different race group covariates. 


```{r cache=T}
formula = violent_crime ~ perc_black + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for violent crime rate model"
             )
```


```{r cache=T}
formula = violent_crime ~ perc_not_visible_minority + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent not a visible minority", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for violent crime rate model"
             )
```

```{r cache=T}
formula = violent_crime ~ perc_asian + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for violent crime rate model"
             )
```

## Theft crime models

The following models have theft crime as the response, and have the same socio-economic covariates, but with different race group covariates. 


```{r cache=T}
formula = theft ~ perc_black + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for theft crime rate model"
             )
```

```{r cache=T}
formula = theft ~ perc_not_visible_minority + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent not visible minority", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for theft crime rate model"
             )
```

```{r cache=T}
formula = theft ~ perc_asian + unemp_rate +
  perc_children + perc_youth + perc_seniors +  
  perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for theft crime rate model"
             )
```

## Diversity index models

The following models use diversity index as a covariate. 


```{r}
# Calculate diversity index
races = data %>% dplyr::select(perc_black, perc_not_visible_minority,perc_Arab, perc_lat_american, perc_south_asian, perc_southeast_asian, perc_west_asian, perc_chinese,
         perc_Filipino, , perc_Japanese, perc_Korean) %>% st_drop_geometry()
races = races/100
races$other = 1-rowSums(races)
di = rowSums(races * (1-races))
data$di = di*100
```


```{r cache=T}
formula = violent_crime ~ di + unemp_rate + perc_children + perc_youth + perc_seniors + perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for violent crime rate model, with demographic and diversity index covariates"
             )
```

```{r cache=T}
formula = theft ~ di + unemp_rate + perc_children + perc_youth + perc_seniors +
   perc_low_income + perc_female + perc_postsecondary +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Unemployment rate",
                 "Percent children", "Percent youth", "Percent seniors",
                 "Percent low-income","Percent female",
                 "Percent with \n postsecondary education",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for theft crime rate model, with demographic and diversity index covariates"
             )
```

## No socioeconomic variables models

The following models do not control for socioeconomic variables. 

```{r cache=T}
formula = violent_crime ~ perc_black +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for violent crime rate model, without controlling for socio-economic variables"
             )
```

```{r cache=T}
formula = theft ~ perc_black +
  f(spatial, model = "bym2", graph = g, hyper = prior)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop,
  control.predictor = list(compute = TRUE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="posterior means, medians and 95% CIs for theft crime rate model, without controlling for socio-economic variables"
             )
```
