---
title: "Appendix 2: NYC Data Models"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F, cache=T)
```

```{r message=F}
library(sf)
library(INLA)
library(tidyverse)
library(spdep)
```

```{r include=F}
data_csv = read_csv("data/nyc.csv")
data = st_read("data/nyc.shp")
names(data) = c(names(data_csv), "geometry")
```

```{r}
# transform data
data$pop_white_pct_est = data$pop_white_pct_est * 100
data$pop_black_pct_est = data$pop_black_pct_est * 100
data$pop_hisp_pct_est = data$pop_hisp_pct_est * 100
data$pop_asian_pct_est = data$pop_asian_pct_est * 100
data$pop_ba_above_pct_est = data$pop_ba_above_pct_est * 100
data$pop_inpov_pct_est = data$pop_inpov_pct_est * 100
```

```{r}
# Filter out missing data
data = data %>% drop_na()
```

```{r}
# specify priors
hyperpriors = list(
  prec = list(
    prior = "pc.prec",
    param = c(0.5 / 0.31, 0.01)),
  phi = list(
    prior = "pc",
    param = c(0.5, 0.5))
  )
```

```{r message=F}
# Compute adjacency matrix and graph
nb = poly2nb(data)
nb2INLA("map.adj", nb)
g = inla.read.graph(filename = "map.adj")

# Neighbourhood index 
spatial = 1:nrow(data)
```

## Violent crime models

The following models have reported violent crime as the response, and have the same socio-economic covariates, but with different race group covariates. 

```{r cache=T}
formula = n_violent_crime ~ pop_white_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = as.data.frame(data), 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent White", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=3,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model."
             )
```


```{r cache=T}
formula = n_violent_crime ~ pop_black_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model."
             )
```


```{r cache=T}
formula = n_violent_crime ~ pop_hisp_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Hispanic", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model."
             )
```


```{r cache=T}
formula = n_violent_crime ~ pop_asian_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model."
             )
```


## Theft crime models

The following models have reported theft crime as the response, and have the same socio-economic covariates, but with different race group covariates. 


```{r cache=T}
formula = n_theft_crime ~ pop_white_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = as.data.frame(data), 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent White", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported theft crime rate model."
             )
```



```{r cache=T}
formula = n_theft_crime ~ pop_black_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported theft crime rate model."
             )
```


```{r cache=T}
formula = n_theft_crime ~ pop_hisp_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Hispanic", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported theft crime rate model."
             )
```


```{r cache=T}
formula = n_theft_crime ~ pop_asian_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported theft crime rate model."
             )
```


## Violent arrest models

The following models have violent-crime arrests as the response, and have the same socio-economic covariates, but with different race group covariates. 


```{r cache=T}
formula = n_violent_arrest ~ pop_white_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = as.data.frame(data), 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent White", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model."
             )
```



```{r cache=T}
formula = n_violent_arrest ~ pop_black_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model."
             )
```


```{r cache=T}
formula = n_violent_arrest ~ pop_hisp_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Hispanic", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model."
             )
```


```{r cache=T}
formula = n_violent_arrest ~ pop_asian_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model."
             )
```



## Theft arrest 

The following models have theft arrests as the response, and have the same socio-economic covariates, but with different race group covariates. 


```{r cache=T}
formula = n_theft_arrest ~ pop_white_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = as.data.frame(data), 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent White", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for theft arrest rate model."
             )
```



```{r cache=T}
formula = n_theft_arrest ~ pop_black_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for theft arrest rate model."
             )
```


```{r cache=T}
formula = n_theft_arrest ~ pop_hisp_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Hispanic", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for theft arrest rate model."
             )
```


```{r cache=T}
formula = n_theft_arrest ~ pop_asian_pct_est+pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Asian", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for theft arrest rate model."
             )
```



## Diversity Index Models

The following models use diversity index as a covariate.


```{r}
# Calculate diversity index
races = data %>% dplyr::select(pop_white_pct_est,pop_black_pct_est,pop_hisp_pct_est,pop_asian_pct_est) %>% st_drop_geometry()
races = races/100
races$other = 1-rowSums(races)
di = rowSums(races * (1-races))
data$di = di*100
```

```{r cache=T}
formula = n_violent_arrest ~ di + pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model, with demographics and diversity index covariates."
             )
```



```{r cache=T}
formula = n_theft_arrest ~ di + pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for theft arrest rate model, with demographics and diversity index covariates."
             )
```

```{r cache=T}
formula = n_violent_crime ~ di + pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model, with demographics and diversity index covariates."
             )
```



```{r cache=T}
formula = n_theft_crime ~ di + pop_ba_above_pct_est+pop_inpov_pct_est+
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Diversity Index", 
                 "Percent 25 years or older with \n at least a Bachelor's degree",
                 "Percent with income below \n poverty line",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported theft crime rate model, with demographics and diversity index covariates."
             )
```




## No socio-economic variables model

The following model does not control for socio-economic variables. 

```{r cache=T}
formula = n_violent_arrest ~ pop_black_pct_est +
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for violent arrest rate model, without controlling for socio-economic variables"
             )
```

```{r cache=T}
formula = n_violent_crime ~ pop_black_pct_est +
  f(spatial, model = "bym2", graph = g, hyper = hyperpriors)

res <- inla(formula,
  family = "poisson", data = data, 
  E=pop_total_est,
  control.predictor = list(compute = TRUE),
  control.fixed = list(mean=0, mean.intercept=0, prec = 100^(-1), prec.intercept = 100^(-1)),
  control.inla=list(strategy="laplace", fast=FALSE)
)
```

```{r}
s = summary(res)
tb = rbind(s$fixed[,c(1, 4, 3, 5)], s$hyperpar[,c(1, 4, 3, 5)])
row.names(tb) = c("Intercept", "Percent Black",
                 "Precision",
                 "Phi")
knitr::kable(tb, 
             col.names=c("Mean", "Median", "95% CI Lower Bound", 
                         "95% CI Upper Bound"),
             digits=4,
             caption="Posterior means, medians and 95% credible intervals for reported violent crime rate model, without controlling for socio-economic variables"
             )
```

